\large How resource abundance and stochasticity affect animals' space-use requirements

### 2023-09-20 writing/appendix-2-simulations.Rmd
updated based on Chrirs's comments

---
title: \large How resource abundance and stochasticity affect animals' space-use requirements
subtitle: "Appendix 2: Simulations"
author:
  - name: Stefano Mezzini
    email: stefano.mezzini@ubc.ca
    institute: [biol, braes]
  - name: E. Patrícia Medici
    institute: [tapir, iucn, escas]
  - name: Michael J. Noonan
    email: michael.noonan@ubc.ca
    institute: [biol, braes, cmps]
institute:
  - biol: Department of Biology, The University of British Columbia Okanagan, Kelowna, British Columbia, Canada.
  - braes: Okanagan Institute for Biodiversity, Resilience, and Ecosystem Services, The University of British Columbia Okanagan, Kelowna, British Columbia, Canada.
  - tapir: Lowland Tapir Conservation Initiative (LTCI), Instituto de Pesquisas Ecológicas (IPÊ), Rodovia Dom Pedro I, km 47, Nazaré Paulista, São Paulo 12960-000, Brazil.
  - iucn: IUCN SSC Tapir Specialist Group (TSG), Campo Grande, Brazil.
  - escas: Escola Superior de Conservação Ambiental E Sustentabilidade (ESCAS/IPÊ), Rodovia Dom Pedro I, km 47, Nazaré Paulista, São Paulo 12960-000, Brazil.
  - cmps: Department of Computer Science, Math, Physics, and Statistics, The University of British Columbia Okanagan, Kelowna, British Columbia, Canada.
bibliography: 'hr-resource-stoch.bib'
csl: 'the-american-naturalist.csl'
fontsize: 12pt
header-includes:
    - \usepackage{setspace}\doublespacing # for double-spaced text
    - \usepackage[small]{titlesec} # for smaller font for headings
    - \usepackage{caption} # for more customization of captions
    - \captionsetup[figure]{font=scriptsize} # smaller caption text
    - \usepackage{hanging} # for hanging indents in references
subparagraph: true # needed for \usepackage[small]{titlesec}
output:
  bookdown::pdf_document2:
    toc: false
    pandoc_args:
      - '--lua-filter=scholarly-metadata.lua' # for author affiliations
      - '--lua-filter=author-info-blocks.lua' # for author affiliations
---

@@

ggplot(all) +
  facet_wrap(~ replicates) +
  geom_line(aes(animal, diff_est, group = paste(mean, variance)),
            alpha = 0.1) +
  geom_hline(yintercept = 0, color = 'darkorange') +
  labs(x = 'Time',
       y = 'Difference in average daily exploration time (hours)')
```

```{r, echo=FALSE, eval=TRUE, fig.cap="Difference in average time required to reach satiety for each scenario in figure 1 (indicated by each line) and different amounts of tracks (indicated at the top of each facet). The differences are relative to simulations with $2^{10}=1024$ tracks for each time point on the x axis (and each scenario in figure 1). The amount of time stabilizes for $\\gtrapprox 100$ tracks, which suggests that the home range estimates should also be stable with more than 100 tracks.", fig.pos='!h'}
knitr::include_graphics('../figures/5-by-5-sensitivity-analysis/time-to-satiety-difference.png',
                        rel_path = FALSE)
```

@@

```{r, echo=FALSE, eval=TRUE, fig.cap="Ratios of average time required to reach satiety for each scenario in figure 1 (indicated by each line) and different amounts of tracks (indicated at the top of each facet). The ratios are relative to simulations with $2^{10}=1024$ tracks for each time point on the x axis (and each scenario in fig. 1). The amount of time stabilizes for $\\gtrapprox 100$ tracks, which suggests that the home range estimates should also be stable with more than 100 tracks.", fig.pos='!h'}
knitr::include_graphics('../figures/5-by-5-sensitivity-analysis/time-to-satiety-rel-difference.png',
                        rel_path = FALSE)
```

@@

saveRDS(saturation_days, 'simulations/hr-saturation-days.rds')

@@

```{r, eval=TRUE, echo=FALSE, fig.cap="Estimated space-use requirements as a function of the number of days sampled for an animal in a habitat with the highest $\\e(R)$ and lowest $\\var(R)$ (left) and an animal with the lowest $\\e(R)$ and highest $\\var(R)$ (right). In both cases, 100 days are sufficient to produce stable estimates of space-use requirements.", cache=FALSE}
knitr::include_graphics('../figures/hr-over-days.png', rel_path = FALSE)
```

@@

```{r, eval=TRUE}
# marginal effects of mu and sigma only (linear on the log link scale)
m_1 <- gam(
  list(
    # predictor for the mean
    hr ~ quantile + mu + sigma2,
    # predictor for the scale (variance = scale * mean)
    ~ quantile + mu + sigma2),
  family = gammals(),
  data = sims)

@@

AIC(m_1, m_2, m_3) # smooth model with interactions is best

BIC(m_1, m_2, m_3) # smooth model with interactions is best

m <- m_3 # select the best model
summary(m)
```

@@

preds_mu <- bind_cols( # bind new data and predictions
  newd_mu,
  predict(m, newdata = newd_mu, type = 'link', se.fit = TRUE) %>%
    data.frame() %>% # convert list to data frame
    # 95% CIs assuming Gaussian credible intervals on the link scale
    transmute(hr = exp(fit.1),
              hr_lwr = exp(fit.1 - se.fit.1 * 1.96),
              hr_upr = exp(fit.1 + se.fit.1 * 1.96)))

# effect of sigma2
newd_sigma2 <-
  expand_grid(mu = mean(sims$mu),
              sigma2 = seq(min(sims$sigma2), max(sims$sigma2),
                           length.out = 400),
              quantile = unique(sims$quantile))

@@

# final figure
e_r <- 'Resource abundance'
v_r <- 'Resource stochasticity'

@@

sims_l <- sims %>%
  mutate(e = hr - predict(m, newdata = sims, type = 'response')[, 1]) %>%
  pivot_longer(c(mu, sigma2)) %>%
  mutate(x = if_else(name == 'mu', e_r, v_r))

ggplot() +
  facet_grid(. ~ x, scales = 'free', switch = 'x') + # label below x axis
  geom_point(aes(value, hr), filter(sims_l, quantile == 'hr_95'),
             alpha = 0.1, color = pal[3]) +
  # 95% CIs are present but hidden under the estimates
  geom_ribbon(aes(value, ymin = hr_lwr, ymax = hr_upr, fill = x),
              filter(preds, quantile == 'hr_95'), alpha = 0.5) +
  geom_line(aes(value, hr, color = x),
            filter(preds, quantile == 'hr_95')) +
  scale_color_manual(values = pal) +
  scale_x_continuous(NULL, breaks = NULL) + 
  scale_y_continuous('Home range size', breaks = NULL) +
  theme_bw() +
  # to make facet strip look like the x axis
  theme(legend.position = 'none',
        strip.background = element_blank(),
        strip.text = element_text(size = 12))
```

>>>>

---
title: \large How resource abundance and stochasticity affect animals' space-use requirements
subtitle: "Appendix 2: Simulations"
author:
  - name: Stefano Mezzini
    email: stefano.mezzini@ubc.ca
    institute: [biol, braes]
  - name: Chris H. Fleming
    institute: [ucf, smith]
  - name: E. Patrícia Medici
    institute: [tapir, iucn, escas]
  - name: Michael J. Noonan
    email: michael.noonan@ubc.ca
    institute: [biol, braes, cmps]
institute:
  - biol: Department of Biology, The University of British Columbia Okanagan, Kelowna, British Columbia, Canada.
  - braes: Okanagan Institute for Biodiversity, Resilience, and Ecosystem Services, The University of British Columbia Okanagan, Kelowna, British Columbia, Canada.
  - ucf: Department of Biology, University of Central Florida, Orlando, Florida 32816, United States.
  - smith: Smithsonian Conservation Biology Institute, National Zoological Park, 1500 Remount Rd., Front Royal, VA 22630, United States.
  - tapir: Lowland Tapir Conservation Initiative (LTCI), Instituto de Pesquisas Ecológicas (IPÊ), Rodovia Dom Pedro I, km 47, Nazaré Paulista, São Paulo 12960-000, Brazil.
  - iucn: IUCN SSC Tapir Specialist Group (TSG), Campo Grande, Brazil.
  - escas: Escola Superior de Conservação Ambiental E Sustentabilidade (ESCAS/IPÊ), Rodovia Dom Pedro I, km 47, Nazaré Paulista, São Paulo 12960-000, Brazil.
  - cmps: Department of Computer Science, Math, Physics, and Statistics, The University of British Columbia Okanagan, Kelowna, British Columbia, Canada.
bibliography: 'hr-resource-stoch.bib'
csl: 'the-american-naturalist.csl'
fontsize: 12pt
header-includes:
    - \usepackage{setspace}\doublespacing # for double-spaced text
    - \usepackage[small]{titlesec} # for smaller font for headings
    - \usepackage{caption} # for more customization of captions
    - \captionsetup[figure]{font=scriptsize} # smaller caption text
    - \usepackage{hanging} # for hanging indents in references
subparagraph: true # needed for \usepackage[small]{titlesec}
output:
  bookdown::pdf_document2:
    toc: false
    pandoc_args:
      - '--lua-filter=scholarly-metadata.lua' # for author affiliations
      - '--lua-filter=author-info-blocks.lua' # for author affiliations
---

@@

ggplot(all) +
  facet_wrap(~ replicates) +
  geom_line(aes(animal, diff_est, group = paste(mean, variance)),
            alpha = 0.1) +
  geom_hline(yintercept = 0, color = 'red') +
  labs(x = 'Time',
       y = 'Difference in average daily exploration time (hours)')
```

```{r, echo=FALSE, eval=TRUE, fig.cap="Difference in average time required to reach satiety for each scenario in figure 1 (indicated by each line) and different amounts of tracks (indicated at the top of each facet). The differences are relative to simulations with $2^{10}=1024$ tracks for each time point on the x axis (and each scenario in figure 1). The amount of time stabilizes for $\\gtrapprox 100$ tracks, which suggests that the home range estimates should also be stable with more than 100 tracks.", fig.pos='!h', out.width='\\textwidth'}
knitr::include_graphics('../figures/5-by-5-sensitivity-analysis/time-to-satiety-difference.png',
                        rel_path = FALSE)
```

@@

```{r, echo=FALSE, eval=TRUE, fig.cap="Ratios of average time required to reach satiety for each scenario in figure 1 (indicated by each line) and different amounts of tracks (indicated at the top of each facet). The ratios are relative to simulations with $2^{10}=1024$ tracks for each time point on the x axis (and each scenario in fig. 1). The amount of time stabilizes for $\\gtrapprox 100$ tracks, which suggests that the home range estimates should also be stable with more than 100 tracks.", fig.pos='!h', out.width='\\textwidth'}
knitr::include_graphics('../figures/5-by-5-sensitivity-analysis/time-to-satiety-rel-difference.png',
                        rel_path = FALSE)
```

@@

@@

```{r, eval=TRUE, echo=FALSE, fig.cap="Estimated space-use requirements as a function of the number of days sampled for an animal in a habitat with the highest $\\e(R)$ and lowest $\\var(R)$ (left) and an animal with the lowest $\\e(R)$ and highest $\\var(R)$ (right). In both cases, 100 days are sufficient to produce stable estimates of space-use requirements.", cache=FALSE, out.width='75%', fig.height=8, fig.width=12}
knitr::include_graphics('../figures/hr-over-days.png', rel_path = FALSE)
```

@@

```{r, eval=TRUE, fig.height=8, fig.width=8, out.width='65%'}
# marginal effects of mu and sigma only (linear on the log link scale)
m_1 <- gam(
  list(
    # predictor for the mean
    hr ~ quantile + mu + sigma2,
    # predictor for the scale (variance = scale * mean)
    ~ quantile + mu + sigma2),
  family = gammals(),
  data = sims)

@@

# check if CV = sigma/mu is better than sigma
m_4 <- gam(
  list(
    hr ~ quantile + s(mu) + s(cv) + ti(mu, cv, k = 5),
    ~ quantile + s(mu) + s(cv) + ti(mu, cv, k = 5)),
  family = gammals(),
  data = mutate(sims, cv = sqrt(sigma2) / mu))

# ti term of the scale parameter looks over-fit and too uncertain
plot(m_4, all.terms = TRUE, pages = 1, scheme = 3, scale = 0)

AIC(m_1, m_2, m_3, m_4) # not much of a difference between m_3 and m_4
BIC(m_1, m_2, m_3, m_4) # not much of a difference between m_3 and m_4
```

```{r, eval=TRUE}
# the difference in df (and significance) is in the ti.1() term
summary(m_3)
```

```{r, eval=TRUE}
summary(m_4)
```

```{r, eval=TRUE, fig.height=8, fig.width=8, out.width='65%'}
m <- m_3 # select the best model
```

@@

preds_mu <- bind_cols( # bind new data and predictions
  newd_mu,
  predict(m, newdata = newd_mu, type = 'link', se.fit = TRUE) %>%
    data.frame() %>% # convert list to data frame
    # 95% CIs assuming Gaussian credible intervals on the link scale
    transmute(hr = exp(fit.1),
              hr_lwr = exp(fit.1 - se.fit.1 * 1.96),
              hr_upr = exp(fit.1 + se.fit.1 * 1.96)))
```

```{r, eval=TRUE}
# effect of sigma2
newd_sigma2 <-
  expand_grid(mu = mean(sims$mu),
              sigma2 = seq(min(sims$sigma2), max(sims$sigma2),
                           length.out = 400),
              quantile = unique(sims$quantile))

@@

# final figure (Rmd throws and error with label_bquote)
e_r <- 'paste(bold("Resource abundance, E("), bolditalic("R"), bold(")"))'
v_r <-
  'paste(bold("Resource stochasticity, Var("), bolditalic("R"), bold(")"))'

@@

sims_l <- sims %>%
  mutate(e = hr - predict(m, newdata = sims, type = 'response')[, 1]) %>%
  pivot_longer(c(mu, sigma2)) %>%
  mutate(x = if_else(name == 'mu', e_r, v_r))
```

```{r, eval=TRUE}
ggplot() +
  facet_grid(. ~ x, scales = 'free', switch = 'x',
             labeller = label_parsed) + # label below x axis
  geom_point(aes(value, hr), filter(sims_l, quantile == 'hr_95'),
             alpha = 0.1, color = pal[3]) +
  # 95% CIs are present but hidden under the estimates
  geom_ribbon(aes(value, ymin = hr_lwr, ymax = hr_upr, fill = x),
              filter(preds, quantile == 'hr_95'), alpha = 0.5) +
  geom_line(aes(value, hr, color = x),
            filter(preds, quantile == 'hr_95'), linewidth = 2) +
  scale_color_manual(values = pal) +
  scale_x_continuous(NULL, breaks = NULL) + 
  scale_y_continuous(bquote(paste(bold('Space-use requirements, '),
                                  bolditalic('H'))), breaks = NULL) +
  theme_bw() +
  # to make facet strip look like the x axis
  theme(legend.position = 'none',
        strip.background = element_blank(),
        strip.text = element_text(size = 12))
```

### 2023-09-20 writing/appendix-3-empirical-modeling.Rmd
updated based on Chrirs's comments

---
title: \large How resource abundance and stochasticity affect animals' space-use requirements
subtitle: "Appendix 3: Empirical modeling"
author:
  - name: Stefano Mezzini
    email: stefano.mezzini@ubc.ca
    institute: [biol, braes]
  - name: E. Patrícia Medici
    institute: [tapir, iucn, escas]
  - name: Michael J. Noonan
    email: michael.noonan@ubc.ca
    institute: [biol, braes, cmps]
institute:
  - biol: Department of Biology, The University of British Columbia Okanagan, Kelowna, British Columbia, Canada.
  - braes: Okanagan Institute for Biodiversity, Resilience, and Ecosystem Services, The University of British Columbia Okanagan, Kelowna, British Columbia, Canada.
  - tapir: Lowland Tapir Conservation Initiative (LTCI), Instituto de Pesquisas Ecológicas (IPÊ), Rodovia Dom Pedro I, km 47, Nazaré Paulista, São Paulo 12960-000, Brazil.
  - iucn: IUCN SSC Tapir Specialist Group (TSG), Campo Grande, Brazil.
  - escas: Escola Superior de Conservação Ambiental E Sustentabilidade (ESCAS/IPÊ), Rodovia Dom Pedro I, km 47, Nazaré Paulista, São Paulo 12960-000, Brazil.
  - cmps: Department of Computer Science, Math, Physics, and Statistics, The University of British Columbia Okanagan, Kelowna, British Columbia, Canada.
bibliography: 'hr-resource-stoch.bib'
csl: 'the-american-naturalist.csl'
fontsize: 12pt
header-includes:
    - \usepackage{setspace}\doublespacing # for double-spaced text
    - \usepackage[small]{titlesec} # for smaller font for headings
    - \usepackage{caption} # for more customization of captions
    - \captionsetup[figure]{font=scriptsize} # smaller caption text
    -  \usepackage{hanging} # for hanging indents in references
subparagraph: true # needed for \usepackage[small]{titlesec}
output:
  bookdown::pdf_document2:
    pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
    toc: false
---

@@

knitr::opts_chunk$set(eval = FALSE,         # do not run R code
                      echo = TRUE,          # print R code
                      message = FALSE,      # hide messages
                      comment = '',         # don't add anything before outputs
                      fig.align = 'center', # change figure alignment
                      cache = TRUE,         # cache all chunk outputs
                      singlespacing = TRUE) # use single spacing for code chunks
knitr::opts_knit$set(root.dir = '..')
```

@@

```{r tapir-mw, fig.cap="Seven-day home-range size of a lowland tapir (\\emph{Tapirus terrestris}) in response to changes in mean and variance in resource abundance. (a.) Mean resource abundance, $\\mu(t)$, esimated as the mean NDVI at the locations visited by the tapir. (b.) Varince in resource abundance, $\\sigma^2(t)$, esimated as the average variance in NDVI at the locations visited by the tapir. (c.) Estimated seven-day home range based on the 95\\% utilization quantiles. (d., e.) estimated marginal effects of $\\mu(t)$ and $\\sigma^2(t)$ on home-range size. The model accounted for the marginal effects of $\\mu(t)$ and $\\sigma^2(t)$ and their interaction on mean space-use requirements and the variance around them. (f.) The effect of $\\mu(t)$ does not follow the data closely because $\\e(R)$ and $\\var(R)$ are highly correlated. Consequently, while estimating the effects of $\\e(R)$ and $\\var(R)$ via separate models would allow result in a closer fit, the estimated effects would be inappropriate because they do not disentangle the effects of $\\e(R)$ and $\\var(R)$. See Appendix 3 for additional information. The tapir movement data corresponds to the individual named ``Anna\" from the Cerrado sample of Medici $et~al.$ (2022).", out.width='100%', eval=TRUE, echo=FALSE}
knitr::include_graphics('H:/GitHub/hr-resource-stoch/figures/tapir-example-with-data.png',
                        rel_path = FALSE)
```

@@

# attach the beta location-scale family written by Simon Wood
source('functions/betals.r')

# source all necessary scripts
source('analysis/figures/default-figure-styling.R') # for color palettes
source('earthdata-login-info.R') # personal login info for EarthData
source('functions/window_hr.R') # function to calculate HRs
```

@@

```{r}
# import tapir data from https://github.com/StefanoMezzini/tapirs
anna <- readRDS('../tapirs/models/tapirs-final.rds') %>%
  filter(name.short == 'ANNA')
anna_tel <- anna$data[[1]] # telemetry data

@@

# calculate the 7-day home range estimate
window_hr(
  tel = anna_tel,
  window = 7 %#% 'day', # 1 week of data for sufficient sample size
  dt = 1 %#% 'day', # move window over by a single day each time
  fig_path = 'figures',
  rds_path = 'models')
anna_mw <- readRDS('models/tapirs/CE_31_ANNA-window-7-days-dt-1-days.rds')
anna_mw
```

```{r, echo=FALSE, eval=TRUE}
anna <- readRDS('../tapirs/models/tapirs-final.rds') %>%
  filter(name.short == 'ANNA')
anna_tel <- anna$data[[1]] # telemetry data

@@

anna_mw <- readRDS('models/tapirs/CE_31_ANNA-window-7-days-dt-1-days.rds')
anna_mw
```

@@

```{r mw, eval=TRUE, echo=FALSE, fig.cap="Exploratory figure created by the $\\tt{window\\_hr()}$ function. Panel $\\mathbf{a.}$ shows the tapir's GPS locations, while panel $\\mathbf{b.}$ shows the seven-day home range estimates (95\\% and 50\\% utilization quantiles) with 95\\% confidence intervals.", out.width='\\textwidth'}
knitr::include_graphics('H:/GitHub/hr-resource-stoch/figures/tapirs/CE_31_ANNA-window-7-days-dt-1-days.png', rel_path = FALSE)
```

@@

```{r, eval=TRUE, fig.height=4, fig.width=12, out.width='\\linewidth'}
theme_set(theme_bw()) # need to set the theme again for some reason

# axis labels (spaces in hr_lab to avoid overlap with the plot label)
e_r <- 'Resource abundance'
v_r <- 'Resource stochasticity'
hr_lab <- expression('7-day'~home~range~size~(km^2)~'   ')

l_grobs <-
  lapply(
    list(
      # mean
      ggplot(tapir, aes(date, mu)) +
        geom_line(color = pal[1]) +
        labs(x = NULL, y = e_r),
      # variance
      ggplot(tapir, aes(date, sigma2)) +
        geom_line(color = pal[2]) +
        labs(x = NULL, y = v_r),
      # 95% home range
      ggplot(tapir, aes(date, hr_est_95)) +
        geom_line(color = pal[3]) +
        labs(x = NULL, y = hr_lab)),
    as_grob) # convert to grid graphical objects (grobs)

# align left margins of all plots
aligned_widths <- align_margin(map(l_grobs,
                                   function(x) {x$widths}), 'first')

# Setting the dimensions of plots to the aligned dimensions
for (i in seq_along(l_grobs)) {
  l_grobs[[i]]$widths <- aligned_widths[[i]]
}

@@

```{r, eval=TRUE, fig.height=4, fig.width=12, out.width='\\linewidth'}
theme_set(theme_bw()) # need to set the theme again for some reason

# predict marginal effects for each term
preds <- tibble(mu = gratia:::seq_min_max(tapir$mu, n = 250),
                sigma2 = gratia:::seq_min_max(tapir$sigma2, n = 250)) %>%
  bind_cols(
    # predictions for the marginal effect of mu
    predict(m, newdata = ., terms = 's(mu)', type = 'link',
            se.fit = TRUE) %>%
      as.data.frame() %>% # convert list to data frame
      transmute(hr_mu_est = exp(fit), # transform to the response scale
                hr_mu_lwr = exp(fit - 1.96 * se.fit), # 95% cred. intervals
                hr_mu_upr = exp(fit + 1.96 * se.fit)),
    # predictions for the marginal effect of sigma2
    predict(m, newdata = ., terms = 's(sigma2)', type = 'link',
            se.fit = TRUE) %>%
      as.data.frame() %>%
      transmute(hr_sigma2_est = exp(fit),
                hr_sigma2_lwr = exp(fit - 1.96 * se.fit),
                hr_sigma2_upr = exp(fit + 1.96 * se.fit)))

@@

# align right margins of all plots
aligned_widths <- align_margin(map(r_grobs,
                                   function(x) {x$widths}), 'first')

# Setting the dimensions of plots to the aligned dimensions
for (i in seq_along(r_grobs)) {
  r_grobs[[i]]$widths <- aligned_widths[[i]]
}

>>>>

---
title: \large How resource abundance and stochasticity affect animals' space-use requirements
subtitle: "Appendix 3: Empirical modeling"
author:
  - name: Stefano Mezzini
    email: stefano.mezzini@ubc.ca
    institute: [biol, braes]
  - name: Chris H. Fleming
    institute: [ucf, smith]
  - name: E. Patrícia Medici
    institute: [tapir, iucn, escas]
  - name: Michael J. Noonan
    email: michael.noonan@ubc.ca
    institute: [biol, braes, cmps]
institute:
  - biol: Department of Biology, The University of British Columbia Okanagan, Kelowna, British Columbia, Canada.
  - braes: Okanagan Institute for Biodiversity, Resilience, and Ecosystem Services, The University of British Columbia Okanagan, Kelowna, British Columbia, Canada.
  - ucf: Department of Biology, University of Central Florida, Orlando, Florida 32816, United States.
  - smith: Smithsonian Conservation Biology Institute, National Zoological Park, 1500 Remount Rd., Front Royal, VA 22630, United States.
  - tapir: Lowland Tapir Conservation Initiative (LTCI), Instituto de Pesquisas Ecológicas (IPÊ), Rodovia Dom Pedro I, km 47, Nazaré Paulista, São Paulo 12960-000, Brazil.
  - iucn: IUCN SSC Tapir Specialist Group (TSG), Campo Grande, Brazil.
  - escas: Escola Superior de Conservação Ambiental E Sustentabilidade (ESCAS/IPÊ), Rodovia Dom Pedro I, km 47, Nazaré Paulista, São Paulo 12960-000, Brazil.
  - cmps: Department of Computer Science, Math, Physics, and Statistics, The University of British Columbia Okanagan, Kelowna, British Columbia, Canada.
bibliography: 'hr-resource-stoch.bib'
csl: 'the-american-naturalist.csl'
fontsize: 12pt
header-includes:
    - \usepackage{setspace}\doublespacing # for double-spaced text
    - \usepackage[small]{titlesec} # for smaller font for headings
    - \usepackage{caption} # for more customization of captions
    - \captionsetup[figure]{font=scriptsize} # smaller caption text
    -  \usepackage{hanging} # for hanging indents in references
subparagraph: true # needed for \usepackage[small]{titlesec}
output:
  bookdown::pdf_document2:
    pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
    toc: false
---

@@

knitr::opts_chunk$set(eval = FALSE,         # do not run R code
                      echo = TRUE,          # print R code
                      message = FALSE,      # hide messages
                      comment = '',         # don't add anything before outputs
                      fig.align = 'center', # change figure alignment
                      cache = TRUE,         # cache all chunk outputs
                      singlespacing = TRUE) # use single spacing for code chunks
```

@@

```{r tapir-mw, fig.cap="Seven-day home-range size of a lowland tapir (\\emph{Tapirus terrestris}) in response to changes in mean and variance in resource abundance. (a.) Mean resource abundance, $\\mu(t)$, esimated as the mean NDVI at the locations visited by the tapir. (b.) Varince in resource abundance, $\\sigma^2(t)$, esimated as the average variance in NDVI at the locations visited by the tapir. (c.) Estimated seven-day home range based on the 95\\% utilization quantiles. (d., e.) estimated marginal effects of $\\mu(t)$ and $\\sigma^2(t)$ on home-range size. The model accounted for the marginal effects of $\\mu(t)$ and $\\sigma^2(t)$ and their interaction on mean space-use requirements and the variance around them. (f.) The effect of $\\mu(t)$ does not follow the data closely because $\\e(R)$ and $\\var(R)$ are highly correlated. Consequently, while estimating the effects of $\\e(R)$ and $\\var(R)$ via separate models would allow result in a closer fit, the estimated effects would be inappropriate because they do not disentangle the effects of $\\e(R)$ and $\\var(R)$. See Appendix 3 for additional information. The tapir movement data corresponds to the individual named ``Anna\" from the Cerrado sample of Medici $et~al.$ (2022).", out.width='100%', eval=TRUE, echo=FALSE}
knitr::include_graphics('../figures/tapir-example-with-data.png', rel_path = FALSE)
```

@@

# attach the beta location-scale family written by Simon Wood
source('../functions/betals.r')

# source all necessary scripts
source('../analysis/figures/default-figure-styling.R') # for color palettes
source('../earthdata-login-info.R') # personal login info for EarthData
source('../functions/window_hr.R') # function to calculate HRs
```

@@

```{r}
# import tapir data from https://github.com/StefanoMezzini/tapirs
anna <- readRDS('../../tapirs/models/tapirs-final.rds') %>%
  filter(name.short == 'ANNA')
anna_tel <- anna$data[[1]] # telemetry data

@@

# calculate the 7-day home range estimate
window_hr(
  tel = anna_tel,
  window = 7 %#% 'day', # 1 week of data for sufficient sample size
  dt = 1 %#% 'day', # move window over by a single day each time
  fig_path = 'figures',
  rds_path = 'models')
anna_mw <-readRDS('../models/tapirs/CE_31_ANNA-window-7-days-dt-1-days.rds')
anna_mw
```

```{r, echo=FALSE, eval=TRUE}
anna <- readRDS('../../tapirs/models/tapirs-final.rds') %>%
  filter(name.short == 'ANNA')
anna_tel <- anna$data[[1]] # telemetry data

@@

anna_mw <- readRDS('../models/tapirs/CE_31_ANNA-window-7-days-dt-1-days.rds')
anna_mw
```

@@

```{r mw, eval=TRUE, echo=FALSE, fig.cap="Exploratory figure created by the $\\tt{window\\_hr()}$ function. Panel $\\mathbf{a.}$ shows the tapir's GPS locations, while panel $\\mathbf{b.}$ shows the seven-day home range estimates (95\\% and 50\\% utilization quantiles) with 95\\% confidence intervals.", out.width='\\textwidth'}
knitr::include_graphics('../figures/tapirs/CE_31_ANNA-window-7-days-dt-1-days.png')
```

@@

```{r, eval=TRUE, fig.height=4, fig.width=12, out.width='\\linewidth'}
theme_set(theme_bw()) # need to set the theme again for some reason
# axis labels (spaces in hr_lab to avoid overlap with the plot label)
# unicode characters removed because they knitting errors
e_r <- bquote(paste(bold('Resource abundance, E('), bolditalic(R),
                    bold(')')))
v_r <- bquote(paste(bold('Resource stochasticity, Var('), bolditalic(R),
                    bold(')')))
hr_lab <- bquote(bold(paste('7-day home range size (', km^2, ')   ')))

l_grobs <- lapply(
  list(ggplot(tapir, aes(date, mu)) + # mean
         geom_line(color = pal[1]) +
         labs(x = NULL, y = e_r),
       ggplot(tapir, aes(date, sigma2)) + # variance
         geom_line(color = pal[2]) +
         labs(x = NULL, y = v_r),
       ggplot(tapir, aes(date, hr_est_95)) + # 95% home range
         geom_line(color = pal[3]) +
         labs(x = NULL, y = hr_lab)),
  as_grob) # convert to grid graphical objects (grobs)

# align left margins of all plots
aligned_widths <- align_margin(map(l_grobs, \(x) {x$widths}), 'first')

# Setting the dimensions of plots to the aligned dimensions
for(i in seq_along(l_grobs)) l_grobs[[i]]$widths <- aligned_widths[[i]]

@@

```{r, eval=TRUE, fig.height=4, fig.width=12, out.width='\\linewidth'}
theme_set(theme_bw()) # need to set the theme again for some reason
# predict marginal effects for each term
preds <- tibble(mu = gratia:::seq_min_max(tapir$mu, n = 250),
                sigma2 = gratia:::seq_min_max(tapir$sigma2, n = 250)) %>%
  bind_cols(
    # predictions for the marginal effect of mu
    predict(m, newdata = ., terms = 's(mu)', type = 'link',
            se.fit = TRUE) %>%
      as.data.frame() %>% # convert list to data frame
      transmute(hr_mu_est = exp(fit), # transform to the response scale
                hr_mu_lwr = exp(fit - 1.96 * se.fit), # 95% cred. intervals
                hr_mu_upr = exp(fit + 1.96 * se.fit)),
    # predictions for the marginal effect of sigma2
    predict(m, newdata = ., terms = 's(sigma2)', type = 'link',
            se.fit = TRUE) %>%
      as.data.frame() %>%
      transmute(hr_sigma2_est = exp(fit),
                hr_sigma2_lwr = exp(fit - 1.96 * se.fit),
                hr_sigma2_upr = exp(fit + 1.96 * se.fit)))

@@

# align right margins of all plots
aligned_widths <- align_margin(map(r_grobs, \(x) {x$widths}), 'first')

# Setting the dimensions of plots to the aligned dimensions
for(i in seq_along(r_grobs)) r_grobs[[i]]$widths <- aligned_widths[[i]]
